---
title: "Usage deseqAbstraction"
author: "Per Ludvik Brattås"
date: "11/7/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tutorial on deseqAbstraction
 
### Easy analysis and visualization of featureCount 

#### 1. Setup of workspace

```{r setup workspace, message = FALSE}

# clean environment and workspace
rm(list=ls())
gc()

# get path to directory with featurecount output
path <- "/Users/perludvik/Dropbox (MN)/Per/PhD/Projects/DNAmeth/hNES DNMT1 KO/RNAseq/PairedParam/Quant/"
# filename of the featurecount file
fc.file <- "hg38.NCBI.genesBestRefSeq.exon.primary.s2.txt"
# full pathh
fc.file.path <- paste(path,fc.file,sep = "/")

cat("filepath: ",fc.file.path)

```

#### 2. get the deseqAbstraction package

```{r load deseaAbstraction package, message = FALSE}

# load deseqAbstraction
library(devtools)
install_github("perllb/deseqAbstraction")
library(deseqAbstraction)
library(DESeq2)

```


#### 3. Create deseqAbs object for this analysis
##### - give a name to this data (in this case "DNMT1KO""), and the name of the raw featureCounts output file

```{r, create object}

# make deseqAbs object, give name of the featurecount file
dnmt <- deseqAbs$new("DNMT1KO",fc.file.path)

```

#### 4. Run DESeq pipeline
##### - You can try to do this in one function $fullAuto(). 
##### - Full auto has four main steps:
##### (1) - create DEseq object (access with $deseq)
##### (2) - Do diffex analysis (access with $test)
##### (3) - Do varianceStablizingTransformation (access with $VST)
##### (4) - Do RPKM normalization (access with $rpkm)

##### - However, for this automation, it is critical that your deseqAbs object has
##### (1) - the raw featureCount file in the correct format
##### (2) - the colData data.frame with at least one column called "condition", which contain the grouping of your samples, that will later be used for diffex anaysis

```{r, automation}

# try to run the automated pipe from creating deseq object to normalization (rpkm, vst) and diffex analysis!
dnmt$fullAuto()

# get error: you need to define the colData for creating DESeq object. Must be data.frame with a "condition" column that will define the diffex analysis groups
dnmt$colData <- data.frame(condition=c(rep("CTR",3),rep("KO",3)))
dnmt$fullAuto()

```

#### 5. Visualize data
#### 5.1 PCA and sample-to-sample

```{r}

# plot PCA without labels on points
PCAplotter(dat = dnmt$VST,title = "NCBI top 5000",ntop = 5000,color = dnmt$colData$condition,shape=dnmt$colData$condition)

# plot PCA with labels
PCAplotter(dat = dnmt$VST,title = "NCBI top 5000",ntop = 5000,color = dnmt$colData$condition,shape=dnmt$colData$condition,label = dnmt$colData$condition)

# sample-to-sample distance
sampleToSample(data = dnmt$VST)

```

#### 5.2 maPlot

```{r}

# make maPlot: color genes if they have p-adj and log2fc below/above given values (p and l) 
maPlot(test = dnmt$test,"DNMT1 KO","CTR",l = .5,p = .005)

```

#### 5.3 volcano plot

```{r}


# generate volcanoplot. 
volcanoPlot(test = dnmt$test)

# cut y-axis 
volcanoPlot(test = dnmt$test,max = 100)


```

#### 5.4 heatmaps
```{r}

# plot top significant genes
mostSignificantHeat(data = assay(dnmt$VST),test = dnmt$test)

# plot top significant genes, top 20
mostSignificantHeat(data = assay(dnmt$VST),test = dnmt$test,ntop = 20)

# plot top significant genes, top 20, with annotation of columns
mostSignificantHeat(data = assay(dnmt$VST),test = dnmt$test,ntop = 20,a1 = dnmt$colData$condition,n1 = "Treatment")

## plot top significant genes, top 70, with two annotation of columns
mostSignificantHeat(data = assay(dnmt$VST),test = dnmt$test,ntop = 70,a1 = dnmt$colData$condition,n1 = "Treatment",a2 = colnames(dnmt$rawCounts),n2 = "sampleNames")

# plot most variable genes
mostVariableHeat(data = assay(dnmt$VST))

# plot most variable genes
mostVariableHeat(data = assay(dnmt$VST),ntop = 100,a1 = dnmt$colData$condition,n1 = "Treatment")


```

#### 6. get average expression levels of all genes

```{r}
base <- getAverage(dnmt$deseq)
head(base$Mean)
head(base$SD)

```

#### 7. Get significant genes
```{r}
# Get diffex-data of genes with p-adj < 0.001 and log2FC more/less than 0.2
sign <- getSign(x = dnmt$test,p = .001,l = .2)
up <- sign$up
down <- sign$down

# Get names of genes with p-adj < 0.01 and log2FC more/less than .1
signID <- getSignName(x = dnmt$test,p = .01,l = .2)
upID <- signID$up
downID <- signID$down

```

